---
title: "Preliminary concepts"
author: "ogorodriguez"
date: "2020-05-01"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE, message = FALSE,
  collapse = TRUE,
  comment = "#>")
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyverse))
theme_set(theme_light())

```

```{r}
# Loading the needed packages
pacman::p_load(tidyverse)
```


## Introduction

Purrr is all about iteration.  It uses the `map()` functions as substitute of the Base R `apply` functions since its authors claim the purrr ones are more robust and consistent.

Purrr is a package built to work with lists, as much as dplyr was built to work with data frames.

A **list** is an object used to collect other objects of various classes or types.  A list can contain a vector, a data frame, a single number or even other lists. 

Here's an example of a list with some of the characteristics described above.  

```{r}
my_first_list <- list(my_number = 5,
                      my_vector = c("a", "b", "c"),
                      my_df = tibble(a = 1:3, b = c("q", "b", "z"), c = c("bananas", "are", "so very great")))

my_first_list


```

A **data frame** is a special case of a list.  It is a list of vectors of the same length.  

```{r}
# Check the clasee of my_first_list
my_first_list %>% class()
```

